---
title: "Hough Transform Implementation"
author: "Sophia Luo"
date: "2022-12-04"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(terra)
library(raster)
library(spatstat)
# test comment
spatstat.options("monochrome" = TRUE)
spatstat.options(transparent=FALSE)
```
Line simulations (two implementations)
points_generator_2: uses rho = xcos(theta) + ysin(theta)
```{r}
points_generator_2 <- function(nline=5, npoint=30, noise_sd = 1){
    
    x_coords <- c()
    y_coords <- c()
    rho_li = c()
    theta_li = c()
    
    for(i in 1:nline){
        t = 2*pi*runif(npoint, 0, 1)
        m = runif(1, 0, 15) # Coefficient of x
        b = runif(1, 0, 15)
        # Added c
        
        x <- runif(npoint, 0, 15)
        e <- rnorm(npoint, 0, noise_sd)
        
        y <- m*x + b + e
        
        plot(x, y)
        

        r <- b/(x * cos(t) + y * sin(t))
        #for(j in 1:npoint){
        #  # Calculate rho
        #    r[j] = b[j]/(x[j] * cos(t[j]) + y[j] * sin(t[j])  )
        #}
        
        x_coords <- c(x_coords, x)
        y_coords <- c(y_coords, y)
        rho_li = c(rho_li, r)
        theta_li = c(theta_li, t)
    }
    
    points = data.frame(theta=theta_li, rho=rho_li, x_coords, y_coords)
    
    return(points)
}

random_points_uniform <- function(npoints, xlim = c(-100, 100), ylim = c(-100, 100)) {
  random_x <- runif(npoints, min = xlim[1], max = xlim[2])
  random_y <- runif(npoints, min = ylim[1], max = ylim[2])
    
  # Convert to polar coordinates
  theta <- atan(random_y/random_y)
  rho <- sqrt(random_x^2 + random_y^2)
  
  return(list(x_coords = random_x, y_coords = random_y, theta = theta, rho = rho))
}

# TODO: generate random noise in a separate process
random_lines_polar <- function(nline = 5, npoints = 30, 
                               xlim = c(-100, 100), slopesd = 5,
                               percentnoise = 0.05, plot = TRUE, 
                               n_random_points = 0) {
  noise_size <- percentnoise*(xlim[2] - xlim[1])
  slopes <- rnorm(nline, mean = 0, sd = slopesd); slopes
  #slopes <- runif(nline, min = -maxslope, max = maxslope)
  ints <- runif(nline, min = xlim[1], max = xlim[2])
  
  x_coords <- numeric(nline * npoints)
  y_coords <- x_coords
  index <- 1
  for (i in 1:nline) {
    # Set xmin and xmax so that a line does not exceed the bounds of a square graph
    xmin <- xlim[1]
    xmax <- xlim[2]
    if (slopes[i] == 0) { # Horizontal line
      line_x <- runif(npoints, min = xmin, max = xmax)
      line_y <- rep(ints[i], npoints)
      next
    } else if (slopes[i] > 0) {
      xmin <- max((xlim[1] - ints[i])/slopes[i], xlim[1])
      xmax <- min((xlim[2] - ints[i])/slopes[i], xlim[2])
    } else {
      xmin <- max((xlim[2] - ints[i])/slopes[i], xlim[1])
      xmax <- min((xlim[1] - ints[i])/slopes[i], xlim[2])
    }
    line_x <- runif(npoints, min = xmin, max = xmax)
    line_y <- slopes[i]*line_x + ints[i] + rnorm(npoints, mean = 0, sd = noise_size)
    index_end <- index + (npoints - 1)
    x_coords[index:index_end] <- line_x
    y_coords[index:index_end] <- line_y
    #print(index); print(index_end)
    index <- index_end + 1
  }
  
  if (n_random_points > 0) {
    random_x <- runif(n_random_points, min = xlim[1], max = xlim[2])
    random_y <- runif(n_random_points, min = xlim[1], max = xlim[2])
    
    x_coords <- c(x_coords, random_x)
    y_coords <- c(y_coords, random_y)
  }
  if (plot) plot(x_coords, y_coords, xlim = xlim)
  
  # Convert to polar coordinates
  theta <- atan(y_coords/x_coords)
  rho <- sqrt(x_coords^2 + y_coords^2)
  
  if (plot) plot(theta, rho)
  
  return(list(x_coords = x_coords, y_coords = y_coords, theta = theta, rho = rho, slopes = slopes, ints = ints))
}
```
Point generation and visualization
```{r}
set.seed(10)
```
```{r}
#points2 = points_generator_2(nline = 1, noise_sd = 3)
#plot(points2$theta, points2$rho)

# Interesting experiments:
# - More noise than points
# - Lots of lines
# - Needle in a haystack

xmin <- -100
xmax <- 100

# Generate random lines
points <- random_lines_polar(nline = 3, npoints = 100, slopesd = 1, xlim = c(xmin,xmax), percentnoise = 0.01, n_random_points = 10)

ymin <- min(points$y_coords)
ymax <- max(points$y_coords)

# Optional: generate background noise
rand_points <- random_points_uniform(npoints = 200, xlim = c(xmin, xmax), ylim = c(ymin, ymax))

# Calculate max distance from origin. TODO: I'm not sure this is right
d <- sqrt((ymax - ymin)^2 + (xmax- xmin)^2)
raster_base <- raster(ncol = 9, nrow = 10, xmn = xmin, xmx = xmax, ymn = ymin*1.1, ymx = ymax*1.1)

# rasterize() needs a data frame of x and y points to plot
data_points <- data.frame(x = points[["x_coords"]], y = points[["y_coords"]])
rdata_points <- data.frame(x = rand_points$x_coords, y = rand_points$y_coords)

# Add random points to line points
#data_points <- rbind(data_points, rdata_points)

raster_counts <- rasterize(data_points, raster_base, fun=function(x,...)length(x), background = 0)
#plot(x = points[["x_coords"]], y = points[["y_coords"]])
plot(raster_counts); points(data_points)
```
Transform points into traces of polar curves in Hough space
```{r}
theta_steps <- seq(from = 0, to = (pi), by = (pi)/180)
theta <- rep(theta_steps, nrow(data_points))

cos_sin_theta <- matrix(c(cos(theta_steps), sin(theta_steps)), byrow = T, nrow = 2)

rho <- matrix(c(points[["x_coords"]], points[["y_coords"]]), ncol = 2) %*% cos_sin_theta # dimensions: (npoints * nlines) x length(theta_steps)
rho <- t(rho) # Now rows are values of theta, columns are values of rho
collected_points <- matrix(c(theta, rho), ncol = 2)

# Plotting is slow for large n
#plot(collected_points[,1], collected_points[,2], cex = 0.05, xlab = "theta", ylab = "rho")
```
For manual accuracy checking, enter eyeballed estimates here
```{r}
theta <- c(1.05, 0.6, 0.8)
rho <- c(52, -22, 25) # Errors (true - estimate): +77, +49, +32.2
# Estimated slopes
-cos(theta)/sin(theta)
# Estimated intercepts
rho/sin(theta)

# True slopes
points$slopes
# True parameters
points$ints
```

Count the intersections
```{r}
theta_cell_size <- 3 # width (in degrees) by height (rho)
# Increasing number of rows does not improve accuracy of y-intercept
raster_base_polar <- raster(ncol = 180, nrow = 180,  xmn = 0, xmx = pi, ymn = -d*.75, ymx = d*.75)
data_points_polar <- data.frame(x = collected_points[,1], y = collected_points[,2])
raster_counts_polar <- rasterize(data_points_polar, raster_base_polar, fun=function(x,...)length(x), background = 0)
# There are optional parameters for smoothing
#spplot(raster_counts_polar, xlab = "theta", ylab = "rho", xlim = c(0, pi), ylim = c(ymin*1.1,ymax*1.1)) #, useRaster = T, interpolate = T)
```
Turn votes into linear parameters
```{r}
accumulation_mat <- matrix(values(raster_counts_polar), byrow = TRUE, ncol = ncol(raster_base_polar), nrow = nrow(raster_base_polar))

top_quant <- quantile(accumulation_mat, probs = c(0.999))

threshold <- top_quant
voted <- which(accumulation_mat > threshold, arr.ind = TRUE)

# First column is theta = 0 to theta = 180/theta_cell_size
theta_bins <- seq(0, pi, by = (pi/ncol(accumulation_mat)))
# the first row is maximum rho, last row is minimum rho
rho_bins <- seq(ymax, ymin, by = -(ymax-ymin)/nrow(accumulation_mat))


params <- matrix(rep(0, 4*nrow(voted)), ncol = 4)
for (i in 1:nrow(voted)) {
  # Calculate average theta and rho value from cell in the raster grid
  
  # col number = theta bin number
  theta_bin_i <- voted[i,2]
  #theta_i <- (theta_bins[voted[i,2]] + theta_bins[voted[i,2] + 1])
  theta_i <- mean(theta_steps[theta_steps > theta_bins[theta_bin_i] & theta_steps <= theta_bins[theta_bin_i + 1]])
  
  # row number = rho bin number
  rho_bin_i <- voted[i,1]
  rho_i <- (rho_bins[rho_bin_i] + rho_bins[rho_bin_i + 1])
  
  m <- -cos(theta_i)/sin(theta_i)
  b <- rho_i/sin(theta_i)
  
  params[i,] <- c(m, b, theta_i, rho_i)
}

# Check estimated rho and theta
params[,3:4]
```
Reduce number of lines
```{r}
# Imagine a caterpillar crawling down the list of m and b estimates
# If it encounters a value more than 0.1 greater than the last value, it returns
# the mean of the values between its tail and its head, then moves its tail up
# to its head
tidy_up <- function(param_estimates, m_sep_threshold = (pi/180), b_sep_threshold) {
  if (nrow(param_estimates) <= 1) return(param_estimates)
  
  m <- numeric(); b <- numeric()
  tail_end <- 1; head_end <- 1
  # Assumes input is sorted
  while (head_end < nrow(param_estimates)) {
    # TODO: this may not handle parallel lines well
    if (head_end == nrow(param_estimates) | 
        # Check if the next slope in the list is much different from current slope
        (abs(param_estimates[head_end + 1, 1] - param_estimates[head_end, 1]) > m_sep_threshold) )
        # Check if the next intercept in the list is much different 
        #| (abs(param_estimates[head_end + 1, 2] - param_estimates[head_end, 2]) > b_sep_threshold) )
      {
      m <- c(m, mean(param_estimates[tail_end, 1], param_estimates[head_end, 1]))
      b <- c(b, mean(param_estimates[tail_end, 2], param_estimates[head_end, 2]))
      tail_end <- head_end + 1
    }
    head_end <- head_end + 1
  }
  
  return(list(m = m, b = b))
}

tidied_params <- tidy_up(params[,1:2], pi/180, 2*(ymax-ymin)/nrow(accumulation_mat))
tidied_params
```
Plot estimated lines in cartesian coordinates
```{r}
plot(data_points$x, data_points$y, ylim=c(ymin*1.5, ymax*1.5))
for (i in 1:nrow(params)) {
  abline(params[i, 2], params[i, 1])
}

plot(data_points$x, data_points$y, ylim=c(ymin*1.5, ymax*1.5))
for (i in 1:length(tidied_params[[1]])) {
  abline(tidied_params$b[i], tidied_params$m[i])
}
```
Convert simulated data to image for compatibility with implementations that take images as input
```{r}
#par(mar=rep(0, 4))
par(bg=NA)
random_line_image <- function(nline, npoints, slopesd = 1, xlim = c(xmin,xmax), percentnoise, numxbins, numybins, n_random_points = 0) {
  points <- random_lines_polar(nline = nline, npoints = npoints, slopesd = slopesd, xlim = xlim, percentnoise = percentnoise, plot = TRUE, n_random_points = n_random_points)
  points_df <- data.frame(x = points[["x_coords"]], y = points[["y_coords"]])

  pattern <- ppp(points_df$x, points_df$y, xlim, c(min(points_df$y), max(points_df$y)))
  pattern_image <- as.im(pattern)
  plot(pattern_image, main = NULL, box = FALSE, axes = FALSE, legend = FALSE, show.all = FALSE)
  
  #raster_frame <- raster(ncol = numxbins, nrow = numybins, xmn = xlim[1], xmx = xlim[2], ymn = ymin*1.1, ymx = ymax*1.1)
  #raster_points <- rasterize(points_df, raster_frame, field = 1, background = 0)
  #plot(raster_points, col = grey.colors(10, start = 1, end = 0), axes = FALSE, box = FALSE, useRaster = TRUE, interpolate = TRUE, legend = FALSE)
}

# Save images as png
png()

# 2 lines
# No noise
set.seed(1)
random_line_image(nline = 2, npoints = 50, slopesd = 3, xlim = c(-100, 100), percentnoise = 0, numxbins = 150, numybins = 170)
set.seed(1)
random_line_image(nline = 2, npoints = 50, slopesd = 3, xlim = c(-100, 100), percentnoise = 0, numxbins = 150, numybins = 170)
set.seed(2)
random_line_image(nline = 2, npoints = 50, slopesd = 3, xlim = c(-100, 100), percentnoise = 0, numxbins = 150, numybins = 170)

# Small noise
set.seed(0)
random_line_image(nline = 2, npoints = 50, slopesd = 3, xlim = c(-100, 100), percentnoise = 0.01, numxbins = 150, numybins = 170)
set.seed(1)
random_line_image(nline = 2, npoints = 50, slopesd = 3, xlim = c(-100, 100), percentnoise = 0.01, numxbins = 150, numybins = 170)
set.seed(2)
random_line_image(nline = 2, npoints = 50, slopesd = 3, xlim = c(-100, 100), percentnoise = 0.01, numxbins = 150, numybins = 170)

# 4 lines
# No noise
set.seed(4)
random_line_image(nline = 4, npoints = 50, slopesd = 3, xlim = c(-100, 100), percentnoise = 0, numxbins = 150, numybins = 170)
set.seed(5)
random_line_image(nline = 4, npoints = 50, slopesd = 3, xlim = c(-100, 100), percentnoise = 0, numxbins = 150, numybins = 170)
set.seed(6)
random_line_image(nline = 4, npoints = 50, slopesd = 3, xlim = c(-100, 100), percentnoise = 0, numxbins = 150, numybins = 170)
# small noise
set.seed(4)
random_line_image(nline = 4, npoints = 50, slopesd = 3, xlim = c(-100, 100), percentnoise = 0.01, numxbins = 150, numybins = 170)
set.seed(5)
random_line_image(nline = 4, npoints = 50, slopesd = 3, xlim = c(-100, 100), percentnoise = 0.01, numxbins = 150, numybins = 170)
set.seed(6)
random_line_image(nline = 4, npoints = 50, slopesd = 3, xlim = c(-100, 100), percentnoise = 0.01, numxbins = 150, numybins = 170)
dev.off()
```